\begin{homeworkProblem}
    Consider the optimization problem: 
    \begin{subequations}
        \begin{align}
            &\min_{x\in\mathbb R^3}\  \
            \quad e^{-x_1-x_2-x_3}+(x_1+2x_2+3x_3)^4 \\
            &\text{ s.t.}\qquad x_1-x_2-x_3=1 \\
            &\qquad\quad \ \  x_1^2+x_2^2+x_3^2+x_1x_2+x_2x_3+x_1x_3\leq 2\\
            &\qquad\quad \ \  0\leq  x_1\leq 1 \\
            &\qquad\quad \ \   0\leq  x_2\leq 1  \\
            &\qquad\quad \ \   0\leq  x_3\leq 1  
        \end{align}
    \end{subequations}
    \begin{itemize}
        \item [i)] Prove that $x=[1 \  \ 0 \  \ 0]^T$ is a feasible
        point.

        \item [ii)] Using the Weierstrass theorem, prove that the above
        optimization problem has a solution.

        \item [iii)] Prove that every local minimum of the above
        optimization problem is a global minimum.
    \end{itemize}

    \begin{solution}
        \begin{itemize}
            \item [i)] The point $x=[1 \  \ 0 \  \ 0]^\top$ is feasible if 
                each constraint is satisfied at it. The equality constraint is
                satisfied since
                \[
                    x_1 - x_2 - x_3 = 1 - 0 - 0 = 1.
                \]
                The inequality constraint is satisfied since
                \[
                    x_1^2 + x_2^2 + x_3^2 + x_1x_2 + x_2x_3 + x_1x_3
                    = 1^2 + 0^2 + 0^2 + 1\cdot 0 + 0\cdot 0 + 1\cdot 0
                    = 1 \leq 2.
                \]
                The remaining constraints are satisfied since
                \[
                    0 \leq x_1 = 1 \leq 1, \quad
                    0 \leq x_2 = 0 \leq 1, \quad
                    0 \leq x_3 = 0 \leq 1.
                \]
                Therefore, the point $x=[1 \  \ 0 \  \ 0]^\top$ is feasible.

            \item [ii)] The Weierstrass theorem states that an optimization
                problem with a continuous objective function has a finite 
                solution if the feasible set is closed and bounded. In 
                terms of this optimization problem, if the functions 
                defining the inequality and equality constraints are 
                continuous, then the feasible set is closed. The 
                functions
                \[
                    x_1 - x_2 - x_3 - 1, \quad
                    x_1^2 + x_2^2 + x_3^2 + x_1x_2 + x_2x_3 + x_1x_3, \quad
                    -x_1, \quad x_1 - 1, \quad
                    -x_2, \quad x_2 - 1, \quad
                    -x_3, \quad x_3 - 1
                \]
                are all continuous. Therefore, the feasible set is closed. 
                The feasible set is a subset of the box defined by the
                constraints $0 \leq x_i \leq 1$ for $i=1,2,3$, which is
                bounded. Therefore, the feasible set is also bounded. Since
                the objective function
                \[
                    e^{-x_1-x_2-x_3} + (x_1 + 2x_2 + 3x_3)^4
                \]
                is continuous, the optimization problem has a finite solution.

            \item [iii)] For a convex optimization problem, any local minimum
                is also a global minimum. Since the objective function is
                the sum of a composition of an exponential function with an
                affine transformation ($a^\top x$ with $a = (-1, -1, -1)$) 
                and a composition of a convex polynomial function ($ 
                f''(u) = 12u^2$) with an affine transformation ($b^\top x$ with 
                $b = (1, 2, 3)$), both of which are convex, so the objective 
                function is convex. The inequality constraint can be rewritten 
                as
                \[
                    \begin{bmatrix} x_1 & x_2 & x_3 \end{bmatrix}
                    \begin{bmatrix}
                    1 & \frac{1}{2} & \frac{1}{2} \\
                    \frac{1}{2} & 1 & \frac{1}{2} \\
                    \frac{1}{2} & \frac{1}{2} & 1
                    \end{bmatrix}
                    \begin{bmatrix} x_1 \\ x_2 \\ x_3 \end{bmatrix}
                    \leq 2
                \]
                The matrix in the quadratic form is positive semidefinite since
                its eigenvalues are $\lambda_1 = 2$, $\lambda_2 = 
                \frac{1}{2}$, and $\lambda_3 = \frac{1}{2}$, which are all non-negative.
                Therefore, the function defining the inequality constraint is 
                convex. The equality constraint is affine since it can be 
                written as
                \[
                    \begin{bmatrix} 1 & -1 & -1 \end{bmatrix}
                    \begin{bmatrix} x_1 \\ x_2 \\ x_3 \end{bmatrix} - 1 = 0.
                \]

                The functions defining the last inequality constraints are
                also convex since their Hessians are zero matrices. Since the 
                functions defining the inequality constraints are convex and 
                the function defining the equality constraint is affine, the 
                feasible set is convex. Therefore, the optimization problem is 
                convex, and every local minimum is a global minimum.
        \end{itemize}
    \end{solution}
\end{homeworkProblem}
