\begin{homeworkProblem}
    Answer the following questions about coercive functions:
    \begin{itemize}
        \item [i)] Find all coefficients $\alpha_1,...,\alpha_n$ and
            $\beta_1,...,\beta_n$ for which the function 
            \begin{equation}
                \alpha_1x_1^2+\cdots+\alpha_nx_n^2
                +\cos\left(\beta_1x_1^3+\cdots+\beta_nx_n^3\right)
            \end{equation}
            is coercive.
        \item [ii)] Given a constant $a$, prove that the function
            $x_1^2+x_2^2+ax_1x_2$ is coercive if $|a|<2$ and is not coercive
            if $|a|\geq 2$. 
    \end{itemize}

    \begin{solution}
        \begin{itemize}
            \item [i)] For the function to be coercive, $\lim_{\|x\|\to\infty} 
                f(x) = \infty$ must hold. The cosine term is bounded between -1 
                and 1, so as $\|x\| \to \infty$, so the coefficients $\beta_i$ 
                can take on any value. All coefficients $\alpha_i$ must be 
                positive for the quadratic term to grow without bound as 
                $\|x\| \to \infty$ for any $x$. This is because if any 
                $\alpha_i$ is negative, then along the axis corresponding to 
                that variable, that is, when all other variables are zero, the 
                function will tend to $-\infty$ as that variable tends to 
                $\infty$. Therefore, the function is coercive if and only if 
                $\alpha_i > 0$ for all $i = 1, \ldots, n$.

            \item[ii)] The function $x_1^2 + x_2^2 + ax_1x_2$ can be rewritten
                as 
                \[
                    x_1^2 + x_2^2 + ax_1x_2 = x^\top A x = 
                    \begin{bmatrix} x_1 & x_2 \end{bmatrix}
                    \begin{bmatrix} 1 & \frac{a}{2} \\ \frac{a}{2} & 1 \end{bmatrix}
                    \begin{bmatrix} x_1 \\ x_2 \end{bmatrix}
                \]
                
                The eigenvalues of symmetric matrix $A$ are given by the 
                characteristic polynomial $\det(A - \lambda I) = 0$, which 
                gives eigenvalues $\lambda = 1 \pm \frac{a}{2}$, with minimum 
                eigenvalue of $\lambda_min = 1 - \frac{|a|}{2}$. If $|a| < 
                2$, all eigenvalues are strictly positive, and $A$ is positive 
                definite. The function is then lower bounded by $\lambda_{\min}
                \|x\|^2_2$ since
                \[
                    \begin{split}
                        x^\top A x &= x^\top V \Lambda V^\top x \\
                        &= y^\top \Lambda y \\
                        &= \sum_{i=1}^n \lambda_i y_i^2 \\
                        &\geq \lambda_{\min} \sum_{i=1}^n y_i^2 \\
                        &= \lambda_{\min} \|y\|^2_2 \\
                        &= \lambda_{\min} \|x\|^2_2
                    \end{split}
                \]

                where $V$ is the orthogonal matrix of eigenvectors and
                $\Lambda$ is the diagonal matrix of eigenvalues. Since $V$ is 
                orthogonal, $\|y\|_2 = \|x\|_2$. Therefore, $x^\top A x \geq 
                \lambda_{\min} \|x\|^2_2 \to \infty$ as $\|x\| \to \infty$ and 
                the function is coercive. If $|a| = 2$, $A$ is linearly 
                dependent and the null space consists of more than just the 
                zero vector. Therefore, there exist non-zero vectors $x$ such 
                that $Ax = 0$, and along those directions the function equals 0 
                and does not tend to $\infty$ as $\|x\| \to \infty$, and is not
                coercive. When $|a| > 2$, the minimum eigenvalue is negative, 
                and there exists a direction along its eigenvector for which the 
                function tends to $-\infty$ as $\|x\| \to \infty$, and the 
                function is not coercive. Therefore, the function is coercive 
                if $|a| < 2$ and is not coercive if $|a| \geq 2$.

        \end{itemize}
    \end{solution}

\end{homeworkProblem}